---
title: "Invert Sampling Cleaning"
author: "Chris Kracha"
date: "2023-01-28"
output: html_document
---

This document contains cleaning scripts for data from manual invert sampling conducted from 2018-3/2020 and 4/2021-2023. Tables are reformatted to fit our needs, factors are reordered and relabeled, additional metadata is created, taxa columns are renamed, and all rounds of sampling are combined. Products of this document include long and wide tables, as well as summary tables that make the data more accessible.

```{r setup, include=FALSE}
library(dplyr)
library(janitor)
library(vegan) # for analysis
library("hydroTSM") # for time series, made for hydrology applications
library(lubridate)
library(tidyverse)
```

# Clean tables
## Load in tables
```{r message=FALSE}
inv2018csv <- read_csv("raw_data/invert_sampling_downloads/2018_raw_normalized.csv", skip = 5)
inv2019csv <- read_csv("raw_data/invert_sampling_downloads/2019_normalized.csv", skip = 1)
inv2021_22csv <- read_csv("raw_data/invert_sampling_downloads/2021_raw.csv")
```

## Format tables for each year
```{r}
inv2018f <- inv2018csv %>% clean_names() %>%
  # get relevant columns, including species and WQ
  select(1:42, 47:55, -4) %>%
  # keep only the rows with dates
  filter(!is.na(date_1)) %>%
  # select normalized columns
  select(1:3, ends_with("_number_liter")) %>%
  # remove _number_liter from names
  rename_with(~ gsub("_number_liter", "", .x)) %>%
  # replace species NAs with zeroes
  mutate(across(ostracod:amphipod, ~replace_na(.,0)),
          # format date
         date_1 = dmy(date_1))

inv2019f <- inv2019csv %>% clean_names() %>% 
  # get relevant columns, including species and WQ
  select(1:42, -4, -total_number) %>% 
  # keep only the rows with dates
  filter(!is.na(date_1), !is.na(site_2)) %>%
  # make species numeric
  dplyr::mutate(across(ostracod:clam_shrimp, ~as.numeric(.))) %>%
         # replace species NAs with zeroes
  dplyr::mutate(across(ostracod:clam_shrimp, ~replace_na(.,0))) %>%
         # format date
  dplyr::mutate(date_1 = dmy(date_1))

inv2021f <- inv2021_22csv %>% clean_names() %>%
  # remove unused variables
  select(-person_entering_data, -person_that_sorted_the_sample, 
         -total_number, -unknown, -x57, -comments) %>% 
  # remove unfinished samples
  filter(!row_number() %in% c(7, 24, 26),
         # remove CORE samples (no results)
         sample_type %in% c("FB250", "FB 250")) %>%
  # make species numeric
  dplyr::mutate(across(ostracod:larvae, as.numeric)) %>%
  # replace species NAs with zeroes
  dplyr::mutate(across(ostracod:larvae, ~replace_na(.,0))) %>%
  # normalize species counts by liter
  dplyr::mutate(across(ostracod:larvae, ~.x/7.5)) %>%
  # format date
  mutate(date_on_vial = dmy(date_on_vial))

inv2021f[7, 5] <- (inv2021f[7, 5])/10 # fix probable error ??
```

## Rename columns to match
```{r}
# rename columns to match order_family format
inv2018 <- inv2018f %>% 
  dplyr::rename(date = date_1,
         site = site_2,
         sample_type = sample_type_3,
         # add order to names
         hemiptera_corixidae_boatman = corixidae_boatman,
         diptera_chironomid = chironomid,
         diptera_ceratopogonidae = ceratopogonidae,
         diptera_ephydridae = ephydridae,
         diptera_stratiomyidae = stratiomyidae,
         diptera_culicidae = culicidae,
         arachnida_acari = mile, # misspell
         hemiptera_notonectidae = heteroptera_backswimmer,
         coleoptera_sp = aquatic_coleoptera_sp,
         coleoptera_hydrophilidae = coleoptera_hydrophilidae_biter,
         coleoptera_hydraenidae = coleoptera_hydraenidae_stripes,
         coleoptera_dytiscidae = coleoptera_dystiscidae) # fix spelling

inv2019 <- inv2019f %>% 
  dplyr::rename(date = date_1,
         site = site_2,
         sample_type = sample_type_3,
         ephemeroptera = ephemeroptera_sp, # sp is only category
         arachnida_acari = arachnida_acari_mite,
         coleoptera_hydraenidae = coleoperta_hydraenidae)

inv2021 <- inv2021f %>% 
  dplyr::rename(date = date_on_vial,
         diptera_syrphidae = syrphidae,
         plecoptera = plecoptera_sp, # sp is only category
         diptera_ceratopogonidae = diptera_ceratopohonidae,
         coleoptera_hydraenidae = coleoperta_hydraenidae,
         arachnida_acari = arachnida_acari_mite) %>%
  mutate(ephemeroptera = ephemeroptera + ephemeroptera_sp) %>%
  select(-ephemeroptera_sp)
```

## Combine columns and fix factors
```{r}
# combine rows
inv <- bind_rows(inv2018, inv2019, inv2021) %>%
  # standardize sample_type column
  dplyr::mutate(sample_type = case_when(sample_type %in% c("FB", "FB 250") ~ "FB250", # change FB and FB 250 to FB250
                                 str_detect(sample_type, "CORE") ~ "CORE", # change anything with CORE to just CORE
                                 TRUE ~ sample_type), # keep the rest
         ## standardize site column
         site = case_when(site == "NPB 2" ~ "NPB2",
                          site %in% c("NEV", "NWC") ~ "NEC", # most likely typos
                          site == "CVP1" ~ "VP1",
                          site == "VNVP2" ~ "NVP2", 
                          site == "VNVP4" ~ "NVP4",
                          TRUE ~ site),
         # add site_type NCOS/COPR column
         site_type = case_when(site %in% c("CUL1", "MO1", "NDC", "NEC", "NMC", "NPB", 
                                           "NPB1", "NPB2", "NVBR", "NWP", "PIER", "VBR1") ~ "slough",
                               TRUE ~ "pool"),
         # add seasons column
         season = time2season(date, out.fmt = "seasons"), 
         season = as.factor(case_when(season == "autumm" ~ "fall", # rename factor
                            TRUE ~ season)),
         season = factor(season, c("fall", "winter", "spring", "summer")),
         year = year(date)) %>% # reorder seasons
  
  # reorder columns to make sense
  select(date, year, season, site, site_type, sample_type, ostracod, copepod, amphipod, isopod, cladocera, clam_shrimp, annelida_sp, annelida_oligochaete, annelida_polychaete, nematode, mollusk, gastropod_snail, coleoptera_sp, coleoptera_hydrophilidae, coleoptera_tropiscernus, coleoptera_dytiscidae, coleoptera_dytiscidae_agabus, coleoptera_dytiscidae_liodessus_affinis, coleoptera_dytiscidae_colymbetinae, coleoptera_hydraenidae, coleoptera_hydraenidae_larva_joe, coleoptera_curculionidae_weavil, coleoptera_gyrindae_gyrinini, dipteran_sp, diptera_chironomid, diptera_ceratopogonidae, diptera_ephydridae, diptera_stratiomyidae, diptera_culicidae, diptera_syrphidae, tipulidae_larva, ephemeroptera, plecoptera, trichoptera_sp, hemiptera_corixidae_boatman, hemiptera_notonectidae, hemiptera_mesoveliidae, odonate, collembola_springtail, lepidoptera_crambidae, arachnida_acari, hydrachnidia, larvae, terrestrials)

# get list of columns with no IDs
inv[, 10:length(inv)] %>% select_if(colSums(., na.rm = T) != 0) %>% colnames() %>% 
  setdiff(colnames(inv[, 10:length(inv)]), .)
```

# make time period summary tables for normalization
```{r}
# make season and year sample slough totals
inv_szn_samp <- inv %>% filter(site_type == "slough") %>% group_by(season) %>%
  dplyr::summarize(season_samples = n())

inv_yr_samp <- inv %>% filter(site_type == "slough") %>% group_by(year) %>%
  dplyr::summarize(year_samples = n())

inv_site_samp <- inv %>% filter(site_type == "slough") %>% group_by(site, sample_type) %>%
  dplyr::summarize(site_samples = n())

# join sample totals to table
inv <- inv %>% left_join(inv_szn_samp, by = "season") %>% 
  left_join(inv_yr_samp, by = "year") %>% 
  left_join(inv_site_samp, by = c("site", "sample_type")) %>%
  select(date, year, season, site, site_type, sample_type, season_samples, year_samples, site_samples, everything())

inv %>% write.csv("cleaned_data/invert_sampling_tables/all_invert_data.csv")

rm(inv2018csv, inv2019csv, inv2021_22csv, inv2018f, inv2019f, inv2021f)
```

# Summarise and manipulate data
## Make totals table
```{r}

# very roughly ordered from tolerant to sensitive
invTotal <- inv %>% 
  rowwise() %>%
  dplyr::mutate(flies = sum(c_across(dipteran_sp:diptera_syrphidae), na.rm = T),
         boaters = sum(c_across(hemiptera_corixidae_boatman:hemiptera_mesoveliidae), na.rm = T),
         mites = sum(c_across(arachnida_acari:hydrachnidia), na.rm = T),
         crustaceans = sum(c_across(ostracod:clam_shrimp), na.rm = T),
         worms = sum(c_across(annelida_sp:nematode), na.rm = T),
         mollusks = sum(c_across(mollusk:gastropod_snail), na.rm = T),
         beetles = sum(c_across(coleoptera_sp:coleoptera_gyrindae_gyrinini), na.rm = T),
         springtail = collembola_springtail,
         dragonfly = odonate,
         larvae = sum(c_across(c(tipulidae_larva, larvae)), na.rm = T),
         ept = sum(c_across(ephemeroptera:trichoptera_sp), na.rm = T),
         total = sum(c_across(ostracod:terrestrials), na.rm = T)) %>%
  dplyr::select(date:site_samples, flies, boaters, mites, crustaceans, worms, mollusks, 
         beetles, springtail, dragonfly, larvae, ept, total)

invTotal %>% write.csv("cleaned_data/invert_sampling_tables/all_invert_data_totals.csv")
```

## Make long tables
```{r}
invLong <- inv %>% pivot_longer(10:length(inv), 
                                names_to = "taxa", 
                                values_to = "organisms_L") %>%
  dplyr::mutate(taxa = as.factor(taxa)) %>% na.omit()

invLong %>% write.csv("cleaned_data/invert_sampling_tables/all_invert_data_long.csv")


invTotalLong <- invTotal %>% select(-total) %>% pivot_longer(10:(length(invTotal)-1), 
                                                            names_to = "group", 
                                                            values_to = "organisms_L") %>%
  dplyr::mutate(group = fct_relevel(as.factor(group), c('flies', 'boaters', 'mites', 'crustaceans', 'worms', 
                                                 'mollusks', 'beetles', 'springtail', 'dragonfly', 
                                                 'larvae', 'ept'))) %>% na.omit()

invTotalLong %>% write.csv("cleaned_data/invert_sampling_tables/all_invert_data_totals_long.csv")
```

# Visualisation
## Visualise data coverages
```{r}
invTotalLong %>% filter(site_type == "slough") %>% ggplot(aes(y = organisms_L/season_samples, x = season, fill = season)) + geom_bar(stat = "identity") + ggtitle("Average organisms per sample collected by season, slough sites") + ylab("organisms/L per sample") + labs(fill = "taxa group") + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + guides(fill="none")

invTotalLong %>% filter(site_type == "slough", group != "crustaceans") %>% ggplot(aes(y = organisms_L/season_samples, x = season, fill = group)) + geom_bar(stat = "identity") + ggtitle("Average organisms per sample collected by season, slough sites") + ylab("organisms/L per sample") + labs(fill = "taxa group")

inv %>% ggplot(aes(x = season, fill = site_type)) + geom_bar(stat = "count") +
  ggtitle("Samples collected by season") + ylab("sample count") + labs(fill = "site location")

invLong %>% filter(site_type == "slough") %>% ggplot(aes(fill = sample_type, y = organisms_L/year_samples, x = year(date))) + 
  geom_bar(stat = "identity") + ggtitle("Average organisms per sample collected by year") + 
  ylab("organisms/L per sample") + labs(fill = "sample type")

ggplot(inv, aes(fill = sample_type, x = year(date))) + geom_bar(stat = "count") +
  ggtitle("Samples collected by year") + ylab("sample count") + labs(fill = "sample type")

ggplot(inv, aes(x = date, y = site, color = sample_type)) + geom_point(shape = 15) +
  ggtitle("Samples by site and date") + labs(color = "sample type")
```

## Visualise organisms
```{r}

# Visualize yearly difference
ggplot(invLong) 


# Visualize filter beaker taxa groups from slough
invTotalLong %>% filter(sample_type == "FB250", site_type == "slough") %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by slough site, filter beaker method") + ylab("organisms/L") + labs(fill = "taxa group")

#### Visualize filter beaker organisms from slough
invLong %>% filter(sample_type == "FB250", site_type == "slough") %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = taxa)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by slough site, filter beaker method") + ylab("organisms/L") + labs(fill = "taxa group")

invLong %>% filter(sample_type == "FB250", site_type == "slough") %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = taxa)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by slough site, filter beaker method") + ylab("organisms/L") + labs(fill = "taxa group") + theme(legend.position = "none")
####


invTotalLong %>% filter(sample_type == "FB250", site_type == "slough", group != "crustaceans") %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by slough site, filter beaker method \nNo crustaceans") + ylab("organisms/L") + labs(fill = "taxa group")

invTotalLong %>% filter(sample_type == "FB250", site_type == "slough", !(group %in% c("crustaceans", "boaters"))) %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by slough site, filter beaker method \nNo crustaceans or boaters") + 
  ylab("organisms/L") + labs(fill = "taxa group")

invTotalLong %>% filter(sample_type == "FB250", site_type == "slough", !(group %in% c("crustaceans", "boaters", "flies"))) %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by slough site, filter beaker method \nNo crustaceans, boaters, or flies") + 
  ylab("organisms/L") + labs(fill = "taxa group")

# visualize core organisms from slough
invTotalLong %>% filter(sample_type == "CORE", site_type == "slough") %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by slough site, sediment core method") + ylab("organisms/L") + labs(fill = "taxa group")

invTotalLong %>% filter(sample_type == "CORE", site_type == "slough", group != "crustaceans") %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by slough site, sediment core method \nNo crustaceans") + ylab("organisms/L") + labs(fill = "taxa group")

invTotalLong %>% filter(sample_type == "CORE", site_type == "slough", !(group %in% c("crustaceans", "boaters"))) %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by slough site, sediment core method \nNo crustaceans or boaters") + 
  ylab("organisms/L") + labs(fill = "taxa group")

# visualize filter beaker organisms from vernal pools ### NOT CORRECT BELOW
invTotalLong %>% filter(sample_type == "FB250", site_type == "pool") %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by vernal pool, filter beaker method") + ylab("organisms/L") + labs(fill = "taxa group")

invTotalLong %>% filter(sample_type == "FB250", site_type == "pool", !(group %in% c("crustaceans"))) %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by vernal pool, filter beaker method \nNo crustaceans") + ylab("organisms/L") + labs(fill = "taxa group")

# visualize core organisms from vernal pools
invTotalLong %>% filter(sample_type == "CORE", site_type == "pool") %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by vernal pool, sediment core method") + ylab("organisms/L") + labs(fill = "taxa group")

invTotalLong %>% filter(sample_type == "CORE", site_type == "pool", !(group %in% c("crustaceans"))) %>%
  ggplot(aes(x = site, y = organisms_L/site_samples, fill = group)) + geom_bar(stat = "identity") +
  ggtitle("Taxa group by vernal pool, sediment core method \nNo crustaceans") + ylab("organisms/L") + labs(fill = "taxa group")
```

```{r}
invTotalLong %>% filter(sample_type == "FB250", site_type == "slough") %>% 
  ggplot(aes(x = group, y = organisms_L/season_samples, fill = season)) +
  geom_bar(stat = "identity", position = "dodge")

invTotalLong %>% filter(sample_type == "FB250", site_type == "slough", !group %in% c("crustaceans")) %>% 
  ggplot(aes(x = group, y = organisms_L/season_samples, fill = season)) +
  geom_bar(stat = "identity", position = "dodge") + facet_wrap(~site)
```

# Water quality graphs
```{r}
inv2019f2 <- inv2019csv %>% clean_names() %>% 
  # get relevant columns, including species and WQ
  select(1:42, -4, -total_number, 49:56) %>% 
  # keep only the rows with dates
  filter(!is.na(date_1), !is.na(site_2)) %>%
  # make species numeric
  dplyr::mutate(across(ostracod:clam_shrimp, ~as.numeric(.))) %>%
         # replace species NAs with zeroes
  dplyr::mutate(across(ostracod:clam_shrimp, ~replace_na(.,0))) %>%
         # format date
  dplyr::mutate(date_1 = dmy(date_1))

inv2019f2 %>% filter(sample_type_3 == "FB250")

names2019 <- names(inv2019f2[,4:40])

salPlot <- list()
wqTable <- inv2019f2 %>% filter(sample_type_3 == "FB250")
for (i in 1:length(names2019)) {
  salPlot[[i]] <- ggplot(wqTable) + geom_point(aes_string(x = salinity_ppt, y = names2019[[i]])) +
    ggtitle(names2019[[i]])
  
  print(salPlot[[i]])
}

wqTable <- inv2019f2 %>% filter(sample_type_3 == "FB250") %>% mutate(across(ostracod:p_h, as.numeric))

ggplot(wqTable, aes(x = salinity_ppt, y = log(ostracod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("ostracod") + theme_bw()

library(mgcv)
gam_sal_ost <- gam(log(ostracod+1) ~ s(salinity_ppt, bs = "cs"), data = wqTable)
summary(gam_sal_ost)

ggplot(wqTable, aes(x = salinity_ppt, y = log(ostracod+1))) + geom_point() +
    geom_smooth(method = "lm") + ggtitle("ostracod") + theme_bw()

lm_sal_ost <- lm(log(ostracod+1) ~ salinity_ppt, data = wqTable)
summary(lm_sal_ost)

ggplot(wqTable, aes(x = salinity_ppt, y = log(copepod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("copepod")


ggplot(wqTable, aes(x = salinity_ppt, y = log(hemiptera_corixidae_boatman+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("hemiptera_corixidae_boatman")


ggplot(wqTable, aes(x = salinity_ppt, y = log(cladocera+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("cladocera")


ggplot(wqTable, aes(x = salinity_ppt, y = log(diptera_chironomid+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("diptera_chironomid")


ggplot(wqTable, aes(x = salinity_ppt, y = log(diptera_culicidae+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("diptera_culicidae")


ggplot(wqTable, aes(x = salinity_ppt, y = log(amphipod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("amphipod")


ggplot(wqTable, aes(x = salinity_ppt, y = log(hemiptera_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("hemiptera_total")

ggplot(wqTable, aes(x = salinity_ppt, y = log(annelida_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("annelida_total")

ggplot(wqTable, aes(x = salinity_ppt, y = log(odonate+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("odonate")

ggplot(wqTable, aes(x = salinity_ppt, y = log(coleoptera_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("coleoptera_total")

ggplot(wqTable, aes(x = salinity_ppt, y = log(arachnida_acari_mite+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("arachnida_acari_mite")

ggplot(wqTable, aes(x = salinity_ppt, y = log(collembola_springtail+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("collembola_springtail")

ggplot(wqTable, aes(x = salinity_ppt, y = log(gastropod_snail+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("gastropod_snail")

inv %>% count(site)
```

```{r}
ggplot(wqTable, aes(x = p_h, y = log(ostracod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("ostracod") + theme_bw()

library(mgcv)
gam_ph_ost <- gam(log(ostracod+1) ~ s(p_h, bs = "cs"), data = wqTable)
summary(gam_ph_ost)


ggplot(wqTable, aes(x = p_h, y = log(copepod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("copepod")


ggplot(wqTable, aes(x = p_h, y = log(hemiptera_corixidae_boatman+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("hemiptera_corixidae_boatman")


ggplot(wqTable, aes(x = p_h, y = log(cladocera+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("cladocera")


ggplot(wqTable, aes(x = p_h, y = log(diptera_chironomid+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("diptera_chironomid")


ggplot(wqTable, aes(x = p_h, y = log(diptera_culicidae+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("diptera_culicidae")


ggplot(wqTable, aes(x = p_h, y = log(amphipod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("amphipod")


ggplot(wqTable, aes(x = p_h, y = log(hemiptera_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("hemiptera_total")

ggplot(wqTable, aes(x = p_h, y = log(annelida_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("annelida_total")

ggplot(wqTable, aes(x = p_h, y = log(odonate+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("odonate")

ggplot(wqTable, aes(x = p_h, y = log(coleoptera_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("coleoptera_total")

ggplot(wqTable, aes(x = p_h, y = log(arachnida_acari_mite+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("arachnida_acari_mite")

ggplot(wqTable, aes(x = p_h, y = log(collembola_springtail+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("collembola_springtail")

ggplot(wqTable, aes(x = p_h, y = log(gastropod_snail+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("gastropod_snail")
```

```{r}
ggplot(wqTable, aes(x = p_h, y = log(ostracod+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("ostracod")


ggplot(wqTable, aes(x = p_h, y = log(copepod+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("copepod")


ggplot(wqTable, aes(x = p_h, y = log(hemiptera_corixidae_boatman+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("hemiptera_corixidae_boatman")


ggplot(wqTable, aes(x = p_h, y = log(cladocera+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("cladocera")


ggplot(wqTable, aes(x = p_h, y = log(diptera_chironomid+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("diptera_chironomid")


ggplot(wqTable, aes(x = p_h, y = log(diptera_culicidae+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("diptera_culicidae")


ggplot(wqTable, aes(x = p_h, y = log(amphipod+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("amphipod")


ggplot(wqTable, aes(x = p_h, y = log(hemiptera_total+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("hemiptera_total")

ggplot(wqTable, aes(x = p_h, y = log(annelida_total+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("annelida_total")

ggplot(wqTable, aes(x = p_h, y = log(odonate+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("odonate")

ggplot(wqTable, aes(x = p_h, y = log(coleoptera_total+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("coleoptera_total")

ggplot(wqTable, aes(x = p_h, y = log(arachnida_acari_mite+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("arachnida_acari_mite")

ggplot(wqTable, aes(x = p_h, y = log(collembola_springtail+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("collembola_springtail")

ggplot(wqTable, aes(x = p_h, y = log(gastropod_snail+1))) + geom_point() +
    geom_smooth(method = "glm") + ggtitle("gastropod_snail")
```

```{r}
ggplot(wqTable, aes(x = temperature_c, y = log(ostracod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("ostracod")


ggplot(wqTable, aes(x = temperature_c, y = log(copepod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("copepod")


ggplot(wqTable, aes(x = temperature_c, y = log(hemiptera_corixidae_boatman+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("hemiptera_corixidae_boatman")


ggplot(wqTable, aes(x = temperature_c, y = log(cladocera+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("cladocera")


ggplot(wqTable, aes(x = temperature_c, y = log(diptera_chironomid+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("diptera_chironomid")


ggplot(wqTable, aes(x = temperature_c, y = log(diptera_culicidae+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("diptera_culicidae")


ggplot(wqTable, aes(x = temperature_c, y = log(amphipod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("amphipod")


ggplot(wqTable, aes(x = temperature_c, y = log(hemiptera_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("hemiptera_total")

ggplot(wqTable, aes(x = temperature_c, y = log(annelida_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("annelida_total")

ggplot(wqTable, aes(x = temperature_c, y = log(odonate+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("odonate")

ggplot(wqTable, aes(x = temperature_c, y = log(coleoptera_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("coleoptera_total")

ggplot(wqTable, aes(x = temperature_c, y = log(arachnida_acari_mite+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("arachnida_acari_mite")

ggplot(wqTable, aes(x = temperature_c, y = log(collembola_springtail+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("collembola_springtail")

ggplot(wqTable, aes(x = temperature_c, y = log(gastropod_snail+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("gastropod_snail")
```
```{r}
ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(ostracod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("ostracod")


ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(copepod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("copepod")


ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(hemiptera_corixidae_boatman+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("hemiptera_corixidae_boatman")


ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(cladocera+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("cladocera")


ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(diptera_chironomid+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("diptera_chironomid")


ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(diptera_culicidae+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("diptera_culicidae")


ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(amphipod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("amphipod")


ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(hemiptera_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("hemiptera_total")

ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(annelida_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("annelida_total")

ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(odonate+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("odonate")

ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(coleoptera_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("coleoptera_total")

ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(arachnida_acari_mite+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("arachnida_acari_mite")

ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(collembola_springtail+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("collembola_springtail")

ggplot(wqTable, aes(x = conductivity_specific_m_s_cm, y = log(gastropod_snail+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("gastropod_snail")
```

```{r}
ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(ostracod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("ostracod")


ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(copepod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("copepod")


ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(hemiptera_corixidae_boatman+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("hemiptera_corixidae_boatman")


ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(cladocera+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("cladocera")


ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(diptera_chironomid+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("diptera_chironomid")


ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(diptera_culicidae+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("diptera_culicidae")


ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(amphipod+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("amphipod")


ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(hemiptera_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("hemiptera_total")

ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(annelida_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("annelida_total")

ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(odonate+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("odonate")

ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(coleoptera_total+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("coleoptera_total")

ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(arachnida_acari_mite+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("arachnida_acari_mite")

ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(collembola_springtail+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("collembola_springtail")

ggplot(wqTable, aes(x = atmosph_pressure_mm_hg, y = log(gastropod_snail+1))) + geom_point() +
    geom_smooth(method = "gam") + ggtitle("gastropod_snail")
```


