---
title: "Invert Sampling Cleaning"
author: "Chris Kracha"
date: "2023-01-28"
output: html_document
---

This document contains cleaning scripts for data from manual invert sampling conducted from 2018-3/2020 and 4/2021-2023. Tables are reformatted to fit our needs, factors are reordered and relabeled, additional metadata is created, taxa columns are renamed, and all rounds of sampling are combined. Products of this document include long and wide tables, as well as summary tables that make the data more accessible.

```{r setup, include=FALSE}
library(dplyr)
library(janitor)
#library(vegan) # For analysis
library("hydroTSM") # For time series, made for hydrology applications
library(lubridate)
library(tidyverse)
library(fuzzyjoin)
library(mgcv) # Smooth function

```

# Clean tables

## Load in raw data
```{r message=FALSE}
inv2018csv <- read_csv("raw_data/invert_sampling_downloads/2018_normalized.csv", skip = 1)
inv2019csv <- read_csv("raw_data/invert_sampling_downloads/2019_normalized.csv", skip = 1)
inv2020_22csv <- read_csv("raw_data/invert_sampling_downloads/2021_inverts_raw.csv")
wq2020_22csv <- read_csv("raw_data/invert_sampling_downloads/2021_water_qual.csv")
```

## Format tables for each year

Before 2020, **7.5 liters** of water were filtered for invertebrates. Starting in 2020, the number of scoops of water was increased to **70 liters** in order to capture more rare species. Invertebrate counts are normalized by the number of liters of water or sediment collected.

### 2018 data
Already normalized, includes water quality data
```{r}
inv2018 <- inv2018csv %>% clean_names() %>%
  
  # Keep only the rows with data
  filter(!is.na(salinity_ppt_calculated_1)) %>%
  
  # Get relevant columns, including species and water quality
  select(site_2:clam_shrimp, # invert data
         sample_type_48:p_h, # water quality data
         -x5) %>%
  
  # Remove columns with all NAs
  remove_empty(which = "cols") %>%
  
  # Replace dashes with NAs
  mutate(across(dissolved_oxygen_mg_l:p_h, ~ na_if(., "-")),
         
  # Replace synthetic water quality data with NAs
         across(sample_type_48:p_h, ~ ifelse(sample_type_48 == "Synthetic", NA, .))) %>%
  
  # Remove more unused columns
  select(-sample_type_48, -salinity_ppt) %>%
  
  # Rename variables
  rename(site = site_2,
         date = date_3,
         sample_type = sample_type_4,
         
         # Invert taxa names
         ephemeroptera = ephemeroptera_sp, # sp is only category
         arachnida_acari = arachnida_acari_mite,
         coleoptera_hydraenidae = coleoperta_hydraenidae,
         
         # Water quality variables
         DO_mg_L = dissolved_oxygen_mg_l,
         DO_pct = dissolved_oxygen_percent,
         cond_spec_mS_cm = conductivity_specific_m_s_cm,
         sal_ppt = salinity_ppt_calculated_53,
         temp_c = temperature_c,
         baro_mm_Hg = atmosph_pressure_mm_hg,
         pH = p_h) %>%
  
  # Fix water quality data type
  mutate(across(DO_mg_L:pH, as.numeric),
         across(ostracod:clam_shrimp, as.numeric),
         
  # Format date
         date = dmy(date))
```

### 2019 data
Already normalized, includes water quality data
```{r}
inv2019 <- inv2019csv %>% clean_names() %>% 
  
  # Get relevant columns, including species and water quality
  select(date_1:clam_shrimp, 
         dissolved_oxygen_mg_l:p_h,
         -x4, -total_number) %>% 
  
  # Keep only the rows with dates
  filter(!is.na(date_1), !is.na(site_2)) %>%
  
  # Rename variables
  rename(date = date_1,
         site = site_2,
         sample_type = sample_type_3,
         
         # Invert taxa names
         ephemeroptera = ephemeroptera_sp, # sp is only category
         arachnida_acari = arachnida_acari_mite,
         coleoptera_hydraenidae = coleoperta_hydraenidae,
         
         # Water quality variables
         DO_mg_L = dissolved_oxygen_mg_l,
         DO_pct = dissolved_oxygen_percent,
         cond_spec_mS_cm = conductivity_specific_m_s_cm,
         sal_ppt = salinity_ppt,
         temp_c = temperature_c,
         baro_mm_Hg = atmosph_pressure_mm_hg,
         pH = p_h) %>%
  
  # Fix numeric data type
  mutate(across(ostracod:clam_shrimp, as.numeric),
         across(DO_mg_L:pH, ~ as.numeric(.)),
         
  # Format date
         date = dmy(date))
```

### 2020-2022 data
Not normalized by liters, water quality data is separate
```{r}

## Format invert table
inv2021_insects <- inv2020_22csv %>% clean_names() %>%
  
  # Remove unused variables
  select(-person_entering_data, -person_that_sorted_the_sample, 
         -total_number, -unknown, -x57, -comments) %>% 
  
  # Remove unfinished samples
  filter(!row_number() %in% c(6, 7, 24, 26),
         # remove CORE samples (no results)
         sample_type %in% c("FB250", "FB 250")) %>%
  
  # Make species numeric
  mutate(across(ostracod:larvae, as.numeric),
         
  # Fix date data type
         date_on_vial = dmy(date_on_vial),
         
  # Replace species NAs with zeroes
         across(ostracod:larvae, ~ replace_na(., 0)),
  
  # Normalize species counts by liter
         across(ostracod:larvae, ~ .x/70),
         
  # Combine ephemeroptera columns
         ephemeroptera = ephemeroptera + ephemeroptera_sp,
  
  # Make join columns
         year = lubridate::year(date_on_vial),
         month = month(date_on_vial)) %>%
  
  select(-ephemeroptera_sp) %>%
  
  # Rename variables
  rename(date = date_on_vial,
         diptera_syrphidae = syrphidae,
         plecoptera = plecoptera_sp, # sp is only category
         coleoptera_hydraenidae = coleoperta_hydraenidae,
         arachnida_acari = arachnida_acari_mite)

# Fix data entry error
inv2021_insects[6, 5] <- (inv2021_insects[6, 5])/10


## Format water quality table
inv2021_water <- wq2020_22csv %>% clean_names() %>% 
  
  # Remove unused columns
  select(date, site, dissolved_oxygen_mg_l:p_h) %>%
  
  # Remove bad data
  filter(dissolved_oxygen_mg_l != "over") %>%
  
  # Fix NAs where needed
  mutate(across(dissolved_oxygen_mg_l:p_h, ~ ifelse(.x == "n/a", NA, .)),
  
  # Make variables numeric
         across(dissolved_oxygen_mg_l:p_h, as.numeric),
  
  # Fix date data type
         date = dmy(date)) %>%
  
  # Get sample averages of readings at multiple depths
  group_by(date, site) %>%
  summarize(DO_mg_L = mean(dissolved_oxygen_mg_l, na.rm = T),
            DO_pct = mean(dissolved_oxygen_percent, na.rm = T),
            cond_spec_mS_cm = mean(conductivity_specific_m_s_cm, na.rm = T),
            sal_ppt = mean(salinity_ppt, na.rm = T),
            temp_c = mean(temperature_c, na.rm = T),
            baro_mm_Hg = mean(barometric_pressure_mm_hg, na.rm = T),
            pH = mean(p_h, na.rm = T)) %>%
  
  # Refix NAs
  mutate(across(DO_mg_L:pH, ~ ifelse(.x == "NaN", NA, .)),
         
 # Make join columns
         year = year(date),
         month = month(date)) %>%
  
  rename(date_wq = date)

## Join tables
inv2021 <- left_join(inv2021_insects, inv2021_water, by = c("site", "year", "month")) %>% select(-year, -month, -date_wq)

rm(inv2021_insects, inv2021_water)
```

## Combine columns and fix factors
```{r}
# Combine rows
inv_factors <- bind_rows(inv2018, inv2019, inv2021) %>%
  # standardize sample_type column
  dplyr::mutate(sample_type = case_when(sample_type %in% c("FB", "FB 250") ~ "FB250", # Change FB and FB 250 to FB250
                                 str_detect(sample_type, "CORE") ~ "CORE", # Change anything with CORE to just CORE
                                 TRUE ~ sample_type), # Keep the rest
                
  # Standardize site column
         site = case_when(site == "NPB 2" ~ "NPB2",
                          site %in% c("NEV", "NWC") ~ "NEC", # Most likely typos
                          site == "CVP1" ~ "VP1",
                          site == "VNVP2" ~ "NVP2", 
                          site == "VNVP4" ~ "NVP4",
                          TRUE ~ site),
  
  # Add seasons column
         season = time2season(date, out.fmt = "seasons"), 
         season = as.factor(case_when(season == "autumm" ~ "fall", # Rename factor
                            TRUE ~ season)),
         season = fct_relevel(season, c("winter", "spring", "summer", "fall")), # Reorder seasons
  
  # Reorder sites
         site = factor(site, levels = c("MO1", "PIER", "CUL1", "VBR1", # COPR slough
                                        "NVBR", "NMC", "NEC",  # NCOS slough
                                        "NPB", "NPB1", "NPB2", "NWP", "NDC", # NCOS inlets
                                        "DSP", # Dune swale pond
                                        "VWCB6", "VP1", "NVP2", "NVP4", "VNVP10", "VOW1", # North & west campus vernal pools
                                        "VCC7", "VDSN", "VM3", "VM7", "VMMesa")), # Other vernal pools
  
  # Get year
         year = year(date)) %>%
  
  # Reorder columns to make sense
  select(date, year, season, sample_type, site, ostracod, copepod, amphipod, isopod, cladocera, clam_shrimp, annelida_sp, annelida_oligochaete, annelida_polychaete, nematode, mollusk, gastropod_snail, coleoptera_sp, coleoptera_hydrophilidae, coleoptera_tropiscernus, coleoptera_dytiscidae, coleoptera_dytiscidae_agabus, coleoptera_dytiscidae_liodessus_affinis, coleoptera_dytiscidae_colymbetinae, coleoptera_hydraenidae, coleoptera_curculionidae_weavil, coleoptera_gyrindae_gyrinini, dipteran_sp, diptera_chironomid, diptera_ceratopogonidae, diptera_ephydridae, diptera_stratiomyidae, diptera_culicidae, diptera_syrphidae, tipulidae_larva, ephemeroptera, plecoptera, trichoptera_sp, hemiptera_corixidae_boatman, hemiptera_notonectidae, hemiptera_mesoveliidae, odonate, collembola_springtail, lepidoptera_crambidae, arachnida_acari, hydrachnidia, larvae, terrestrials, DO_mg_L:pH)

# Get list of columns with no IDs
colNoID <- inv_factors[, 6:length(inv_factors)] %>% select_if(colSums(., na.rm = T) != 0) %>% colnames() %>% 
  setdiff(colnames(inv_factors[, 6:length(inv_factors)]), .)  

inv_factors <- inv_factors %>% select(-colNoID)

# check no columns are missing (except totals)
unique(c(setdiff(colnames(inv2021), c(colnames(inv), colNoID)), setdiff(colnames(inv2019), c(colnames(inv), colNoID)), setdiff(colnames(inv2018), c(colnames(inv), colNoID))))
```
# Add metadata

## Add variables for site metadata, breaching, and methods
```{r}
# Create site metadata
site_index <- data.frame(site = unique(inv$site)) %>% 
  
# Full site names
  mutate(site_name = c("Culvert", "Main Channel", "North of Venoco Bridge", "South of Venoco Bridge", "East Channel",
                       "Mouth of Slough", "Phelps Bridge", "Dune Swale Pond", "Devereux Creek", "West Pond", 
                       "NCOS Vernal Pool 2", "NCOS Vernal Pool 4", "West Campus Vernal Pool 6", "Del Sol Vernal Pool N",
                       "COPR Vernal Pool 1", "NCOS Vernal Pool 10", "Camino Corto Vernal Pool 7", 
                       "Manzanita Vernal Pool 7",
                       "Ocean Walk Lane Vernal Pool 1", "More Mesa Vernal Pool", "Manzanita Vernal Pool 3", 
                       "COPR Pier", "Upper Phelps Creek", "Lower Phelps Creek"),
         
# Site_type NCOS/COPR column
         site_type = case_when(site %in% c("CUL1", "MO1", "NDC", "NEC", "NMC", "NPB", 
                                           "NPB1", "NPB2", "NVBR", "NWP", "PIER", "VBR1") ~ "slough",
                               TRUE ~ "pool"),

# Site reserve
         site_reserve = case_when(site %in% c("MO1", "PIER", "CUL1", "VBR1", 
                                              "DSP", "VP1", "VWCB6") ~ "COPR",
                                  site %in% c("NVBR", "NMC", "NEC", "NPB", "NPB1", 
                                              "NPB2", "NDC", "NWP", "NVP2", "NVP4", 
                                              "VNVP10", "VOW1") ~ "NCOS",
                                  TRUE ~ "Other"),

# More detailed site information
         
         site_desc = case_when(site %in% c("MO1", "PIER", "CUL1", "VBR1") ~ "Slough in COPR",
                               site %in% c("NVBR", "NMC", "NEC", "NPB") ~ "Slough in NCOS",
                               site %in% c("NPB1", "NPB2", "NDC") ~ "Slough Inlet",
                               site %in% c("DSP", "NWP") ~ "Pond",
                               site %in% c("VP1", "VWCB6", "NVP2", "NVP4", "VNVP10", 
                                            "VOW1", "VCC7", "VDSN", "VM3", "VM7", "VMMesa") ~ "Vernal Pool",
                                TRUE ~ "Extra"))

# Add site metadata
inv_meta <- inv_factors %>% left_join(site_index, by = c("site")) %>%
  
# Add breaching status
  mutate(breach_status = case_when(date < "2018-03-22" ~ "closed",
                                                date > "2018-03-22" & date < "2018-03-31" ~ "tidal",
                                                date > "2018-03-31" & date < "2018-05-31" ~ "post-breach",
                                                date > "2018-05-22" & date < "2019-01-07" ~ "closed",
                                                date > "2019-01-07" & date < "2019-03-20" ~ "tidal",
                                                date > "2019-03-20" & date < "2019-05-20" ~ "post-breach",
                                                # A breach occurred in March 2020, however no sampling was done
                                                # at this time due to the COVID-19 pandemic
                                                date > "2019-05-20" & date < "2021-12-13" ~ "closed",
                                                date > "2021-12-13" & date < "2022-01-13" ~ "tidal",
                                                date > "2022-01-13" & date < "2022-03-13" ~ "post-breach",
                                                date > "2022-01-13" & date < "2022-12-01" ~ "closed"),
         
# Add FB250 pre-filtration sample volume
                      filter_method = case_when(sample_type == "FB250" & year(date) >= "2020" ~ "70 liters",
                                                sample_type == "FB250" & year(date) < "2020" ~ "7.5 liters"),
                      
# Add total organisms variable
                      total = rowSums(across(ostracod:terrestrials), na.rm = T)) %>%
  
  select(date, year, season, breach_status, 
         site, site_type, site_reserve, site_desc, 
         sample_type, filter_method, everything(), total, DO_mg_L:pH) 

rm(inv2018csv, inv2019csv, inv2020_22csv, wq2020_22csv)
```

## Add sample normalization factors
```{r}
# Make season and year sample slough totals
inv_szn_samp <- inv_meta %>% filter(site_type == "slough") %>% group_by(season, sample_type) %>%
  dplyr::summarize(season_samples = n())

inv_yr_samp <- inv_meta %>% filter(site_type == "slough") %>% group_by(year, sample_type) %>%
  dplyr::summarize(year_samples = n())

inv_site_samp <- inv_meta %>% filter(site_type == "slough") %>% group_by(site, sample_type) %>%
  dplyr::summarize(site_samples = n())

# Join sample totals to table
inv <- inv_meta %>% left_join(inv_szn_samp, by = c("season", "sample_type")) %>% 
  left_join(inv_yr_samp, by = c("year", "sample_type")) %>% 
  left_join(inv_site_samp, by = c("site", "sample_type")) %>%
  select(date, year, season, breach_status, 
         site, site_type, site_reserve, site_desc, 
         sample_type, filter_method, season_samples, year_samples, site_samples, everything(), total, DO_mg_L:pH)
```

```{r}
inv %>% write.csv("cleaned_data/invert_sampling_tables/all_invert_data.csv")

# Make log transformation
inv_log <- inv %>% mutate(across(ostracod:terrestrials, ~ log(.x + 1)))

inv_log %>% write.csv("cleaned_data/invert_sampling_tables/all_invert_data_log.csv")

rm(inv2018, inv2019, inv2021, colNoID, 
   inv_site_samp, inv_szn_samp, inv_yr_samp, inv_factors, inv_meta)
```

# Make long table

```{r}
# All data long table
invLong <- inv %>% pivot_longer(ostracod:terrestrials, 
                                names_to = "taxa", 
                                values_to = "organisms_L") %>%
  
  # Remove blanks
  filter(!is.na(organisms_L)) %>%
  
  # Apply normalizations
  mutate(organisms_L_season = organisms_L/season_samples,
         organisms_L_year = organisms_L/year_samples,
         organisms_L_site = organisms_L/site_samples) %>% 
  select(-season_samples, -year_samples, -site_samples) %>%
  
  # Add log transformations
  mutate(organisms_L_log = log(organisms_L+1),
         organisms_L_season_log = log(organisms_L_season+1),
         organisms_L_year_log = log(organisms_L_year+1),
         organisms_L_site_log = log(organisms_L_site+1),
         
  # Fix faxa data type
         taxa = as.factor(taxa))

invLong %>% write.csv("cleaned_data/invert_sampling_tables/all_invert_data_long.csv")
```

# Extra Unused Code

**Make totals table**

very roughly ordered from tolerant to sensitive
invTotal <- inv %>% rowwise() %>% mutate(group = sum(c_across(:), na.rm = T) %>% select(date:site_samples, total) %>%
  mutate(season = factor(season, c("winter", "spring", "summer", "fall")))

invTotal %>% write.csv("cleaned_data/invert_sampling_tables/all_invert_data_totals.csv")

**Grouped taxa long table**
invTotalLong <- invTotal %>% dplyr::select(-total) %>% pivot_longer(flies:ept, 
                                                            names_to = "group", 
                                                            values_to = "organisms_L") %>%
  dplyr::mutate(group = fct_relevel(as.factor(group), c('flies', 'boaters', 'mites', 'crustaceans', 'worms', 
                                                 'mollusks', 'beetles', 'springtail', 'dragonfly', 
                                                 'larvae', 'ept'))) %>%
  filter(!is.na(organisms_L)) %>%
  # Apply normalizations
  mutate(organisms_L_season = organisms_L/season_samples,
         organisms_L_year = organisms_L/year_samples,
         organisms_L_site = organisms_L/site_samples) %>% 
  select(-season_samples, -year_samples, -site_samples)

invTotalLong %>% write.csv("cleaned_data/invert_sampling_tables/all_invert_data_totals_long.csv")


